Exception Handling Bug Hazards in Android

### 2015-09-27 SPRINGER_JOURNAL/msr_journal.tex
extended paper version

@@

\emph{\textbf{RQ 1:}How often developers deal with the exception handling code in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
\begin{itemize}
\item RQ1.1:\emph{How often developers handle exceptions?}
\item RQ1.2:\emph{How often developers throw exceptions?}
\end{itemize}

Another theme that show us how developers may deal with such bug hazards is
 their knowledge about exception handling best practices presented in Section~\ref{sec:best}.
 This leads us to our second research question:

@@

\emph{\textbf{RQ 2:}Do developers know about exception handling best practices?}

@@

\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2.png}
\caption{Developers expertise in Java and Android development.}
\label{fig:devexpertise}
\end{figure}

@@

In this section, we present our findings per research question. 
We also include direct quotes from developers along the text with the answer identification - which can be traced in our dtabase.

@@

\noindent\emph{\textbf{Q1:How often developers deal with the exception handling code in Android development?}}

@@

\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q3Q4.png}
\caption{How often.}
\label{fig:oftenhandle}
\end{figure}

@@

\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q7Q8.png}
\caption{How often.}
\label{fig:oftenhandle}
\end{figure}

@@

>>>>

%%%%GREY BOX

\usepackage{color}
\definecolor{lightgray}{gray}{0.85}

\newcommand\greybox[1]{%
  \vskip\baselineskip%
  \par\noindent\colorbox{lightgray}{%
    \begin{minipage}{.97\columnwidth}%
    \emph{#1}%
    \end{minipage}%
  }%
  \vskip\baselineskip%
}
%%%% END GREY BOX

@@

\emph{\textbf{RQ 1:} How often developers deal with the exception handling in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
\begin{itemize}
\item RQ1.1:\emph{How often developers handle exceptions?}
\item RQ1.2:\emph{How often developers throw exceptions?}
\end{itemize}

Another issue that show us how developers may deal with such bug hazards is
 their knowledge about exception handling best practices presented in Section~\ref{sec:best}.
 This leads us to our second research question:

@@

\emph{\textbf{RQ 2:} Do Android developers adopt exception handling best practices?}

\emph{\textbf{RQ 4:}  Do Android developers consider implementing the exception handling code of Android apps more difficult than general Java applications?}

@@

\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2_new.png}
\caption{Developers expertise in Java and Android development.}
\label{fig:devexpertise}
\end{figure}

@@

In this section, we present our findings per research question. 
We also include a selection of quotes from the exploratory survey 
respondents. Each respondent has an identification which can be traced in our database.

@@

\noindent\emph{\textbf{Q1:How often developers deal with the exception handling in Android development?}}

@@

\begin{figure} \centering \includegraphics[scale=0.35]{charts/often.png}
\caption{How often developers throw (a) and handle (b) exceptions during app development.}
\label{fig:oftenhandle}
\end{figure}

Developers were also asked whether the exception handling code helped with the development of robust Android applications. Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of them said that if not carefully developed the exception handling code may lead to app crashes. Table XXX presents the top reasons why the exception handling code may or may not help de development of robust apps.

\begin{table}
\scriptsize
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Top 5 Resons} & \bfseries{\#} & \bfseries{\%} \\
\hline
\bfseries{For Positively Affecting Robustness} &   &   \\
Programming logic (java.lang and util)  & 52,0\% & 52,0\%\\
Resources (IO, Memory, Batery)  & 52,0\% & 23,9\% \\
Security  & 52,0\% & 4,1\%\\
Concurrency  & 52,0\% & 2,9\% \\
Backward compatibility  & 52,0\% & 5,5\% \\
\bfseries{For Negatively Affecting Robustness} &   &   \\
Specific Exceptions  & 52,0\% & 4,9\%\\
General (Error, Exception, Runtime)  & 52,0\% & 6,7\%\\
Specific Exceptions  & 52,0\% & 4,9\%\\
General (Error, Exception, Runtime)  & 52,0\% & 6,7\%\\
\hline
\end{tabular}
\caption{Top 5 reasons why exception handling helps or impairs the development of robust apps}
\label{tab:tophundrend}
\end{table}

@@

\begin{figure} \centering \includegraphics[scale=0.35]{charts/bestpractices.png}
\caption{The developers awareness of exception handling best practices in Java (a) and Android (b). }
\label{fig:oftenhandle}
\end{figure}

@@

\greybox{Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background)}

Handle life cycle of activities and fragments.

Q10: What makes exception handling in Android different from exception handling in Java?

The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did

Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background).

Application-Activity-Fragment lifecycle.

You can wrap the main function in java in try/catch in android you can\'t but there are equivalents

Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events. It may be hard to ensure that all this cases are properly handled. The logcat is a helpful way of tracking exceptions when developing and testing on real phones.

Any uncatched exception will crash the app. There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT

The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called
